The \dealii Library, Version 9.1

### 2019-06-05 9.1/paper.tex
Add Daniel Garcia-Sanchez to the author list

@@

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Denis Davydov,
    Marc Fehling,
    Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Ross Maguire Kynch,
    Matthias Maier,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.1, 2019},
}

@@

@@

\author[5]{Timo~Heister}
\affil[5]{Mathematical Sciences,
  O-110 Martin Hall,
  Clemson University,
  Clemson, SC 29634, USA
  and Scientific Computing and Imaging Institute,
  72 S Central Campus Drive, Room 3750
  Salt Lake City, UT 84112.
  {\texttt{heister@sci.utah.edu}}}
%
\author[6]{Luca~Heltai}
\affil[6]{SISSA,
  International School for Advanced Studies,
  Via Bonomea 265,
  34136, Trieste, Italy.
  {\texttt{luca.heltai@sissa.it}}}

\author[7]{Martin~Kronbichler}
\affil[7]{Institute for Computational Mechanics,
  Technical University of Munich,
  Boltzmannstr.~15, 85748 Garching, Germany.
  {\texttt{kronbichler@lnm.mw.tum.de}}}

\author[8]{Ross~Maguire~Kynch}
\affil[8]{Renishaw Plc,
  Miskin, CF72 8XY, Wales, UK.
  {\texttt{rkynch@gmail.com}}}

\author[9]{Matthias~Maier}
\affil[9]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

@@

\author[10]{David Wells}
\affil[10]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\item
  The \texttt{FE\_BernardiRaugel} class implements the non-standard
  Bernardi-Raugel (BR) element that can be used to construct a stable
  velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
  element is an enriched version of the $Q_1^d$ element with added bubble
  functions on each edge (in 2d) or face (in 3d). It addresses the fact
  that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
  larger velocity space), and that the $Q_2^d\times Q_0$ combination is
  stable but sub-optimal since the velocity space is too large relative to
  the pressure space to provide additional accuracy commensurate with the
  cost of the large number of velocity unknowns. The BR space is
  thus intermediate to the $Q_1^d$ and $Q_2^d$ spaces.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dedicated support for symbolic algebra, including algebra differentiation}
\label{subsec:sd}
To complement the automatic differentiation features in \dealii{}, this release sees
the first step towards integrating and supporting a highly performant computer algebra
system (CAS) via the SymEngine library.
\marginpar{Could use a reference to SymEngine}
The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces
to SymEngine and forms the basis of symbolic computations, offering a full set of
overloaded operators and a C++ style interface. This class offers the following
basic functionality:
\begin{itemize}
\item symbolic variable definition,
\item symbolic function definition,
\item expression creation using standard C++ syntax,
\item expression parsing,
\item comparison operations,
\item logical operations,
\item conditional expression construction,
\item differentiation,
\item substitution (partial and complete), and
\item serialization.
\end{itemize}
\dealii{} now also provides an extensive set of math operations, with a syntax mimicking that used
in the C++ standard library. Using the \texttt{Expression} class as a basis, we
have developed a set of functions that can be used to create \dealii{} \texttt{Tensor}s
and \texttt{SymmetricTensor}s of symbolic variables and symbolic functions. This
gives full symbolic tensor algebra support using the pre-existing \texttt{Tensor}
and \texttt{SymmetricTensor} classes and associated functions. We have
also implemented a set of utility functions with the following features:
\begin{itemize}
\item \texttt{differentiate} scalar expressions, tensor and symmetric tensors of expressions
with respect to other scalar expressions, tensor and symmetric tensors of expressions;
\item create symbolic substitution maps;
\item resolve explicit dependencies between expressions; and
\item perform scalar and tensor valued substitution (including conversion from symbolic to
real-valued scalars and tensors).
\end{itemize}

@@

GPU support was significantly extended for the current release:
\begin{itemize}
  \item The \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
    classes can now be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
  \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
    class has gained a second template argument \texttt{MemorySpace} which can
    either be \texttt{Host} or \texttt{CUDA}. In the latter case, the data
    resides in GPU memory. By default, the template parameter is
    \texttt{Host} and the behavior is unchanged compared to previous versions.
    When using CUDA, the ghost exchange can be performed either by first copying
    the relevant data to the host, performing MPI communication, and finally
    move the data to the device or, if CUDA-aware MPI is available, by
    performing MPI communication directly between GPUs.
  \item Constrained degrees of freedom: the matrix-free framework now
    supports constrained degrees of freedom. The implementation is based on the algorithms described in
    \cite{ljungkvist2017}. With this addition, both Dirichlet
    boundary conditions and the constraints arising from adaptively
    refined meshes can be imposed within the matrix-free framework. The only restriction is that for two-dimensional meshes the
    finite element degree must be odd.
    \marginpar{Can we explain this restriction? It seems odd (pun intended)...}
    \marginpar{BT: No, we cannot. I have spent at least two full weeks trying to
    understand why it doesn't work, i.e., gives the wrong result, but I don't know. We decided at the time that
    even with the current limitation this was a useful addition.}
    There is no such restriction in three
    dimensions.
  \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
    matrix-free framework can scale to multiple GPUs by taking
    advantage of MPI. Each MPI process can only use one GPU and therefore, if
    multiple GPUs are available in one node, it is necessary to have as many
    ranks as there are GPUs. Using Nvidia Multi-Process Service (MPS), it is also possible
    for multiple processes to use the same GPU. This can be advantageous if the
    amount of work on one rank is not sufficient to fully utilize a GPU.
\end{itemize}

@@

\marginpar{Missing: performance optimizations, new features,
largest runs now at XYZ (Kronbichler?)}

@@

\marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
  libraries come to mind. Please update.}
\dealii can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{Griewank1996a,adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK \cite{lapack}
\item cuSOLVER \cite{cusolver}
\item cuSPARSE \cite{cusparse}
\item Gmsh \cite{geuzaine2009gmsh}
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{Burstedde2018,p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK \cite{slug}
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

@@

\dealii is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:\\
%
% get this from the changes/*/* files using the command listed in the
% release-tasks paper and remove the authors of this paper
%
% updated 5/12/2019
\marginpar{Remove all other authors of this paper}
\marginpar{Do we want to spell out who the Gingko devs are?}
  Giovanni Alzetta,
  Mathias Anselmann,
  Daniel Appel,
  Alexander Blank,
  Vishal Boddu,
  Benjamin Brands,
  Pi-Yueh Chuang,
  T. Conrad Clevenger,
  Sambit Das,
  the Ginkgo developers,
  Stefano Dominici,
  Nivesh Dommaraju,
  Niklas Fehn,
  Isuru Fernando,
  Andreas Fink,
  Daniel Garcia-Sanchez,
  Rene Gassm{\"o}ller,
  Alexander Grayver,
  Joshua Hanophy,
  Logan Harbour,
  Graham Harper,
  Daniel Jodlbauer,
  Stefan Kaessmair,
  Eldar Khattatov,
  Alexander Knieps,
  Uwe K{\"o}cher,
  Kurt Kremitzki,
  Dustin Kumor,
  Ross Kynch,
  Damien Lebrun-Grandie,
  Jonathan Matthews,
  Stefan Meggendorfer,
  Pratik Nayak,
  Lei Qiao,
  Ce Qin,
  Reza Rastak,
  Roland Richter,
  Alberto Sartori,
  Svenja Schoeder,
  Sebastian Stark,
  Antoni Vidal,
  Jiaxin Wang,
  Yuxiang Wang,
  Zhuoran Wang.

>>>>

\usepackage[normalem]{ulem}

@@

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Denis Davydov,
    Marc Fehling,
    Daniel Garcia-Sanchez,
    Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Ross Maguire Kynch,
    Matthias Maier,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.1, 2019},
}

@@

\author[5]{Daniel Garcia-Sanchez}
\affil[5]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
  Institut des NanoSciences de Paris, F-75005, Paris, France
  {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}

@@

\author[6]{Timo~Heister}
\affil[6]{Mathematical Sciences,
  O-110 Martin Hall,
  Clemson University,
  Clemson, SC 29634, USA
  and Scientific Computing and Imaging Institute,
  72 S Central Campus Drive, Room 3750
  Salt Lake City, UT 84112.
  {\texttt{heister@sci.utah.edu}}}
%
\author[7]{Luca~Heltai}
\affil[7]{SISSA,
  International School for Advanced Studies,
  Via Bonomea 265,
  34136, Trieste, Italy.
  {\texttt{luca.heltai@sissa.it}}}

\author[8]{Martin~Kronbichler}
\affil[8]{Institute for Computational Mechanics,
  Technical University of Munich,
  Boltzmannstr.~15, 85748 Garching, Germany.
  {\texttt{kronbichler@lnm.mw.tum.de}}}

\author[9]{Ross~Maguire~Kynch}
\affil[9]{Renishaw Plc,
  Miskin, CF72 8XY, Wales, UK.
  {\texttt{rkynch@gmail.com}}}

\author[10]{Matthias~Maier}
\affil[10]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

@@

\author[11]{David Wells}
\affil[11]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\item
  The \texttt{FE\_BernardiRaugel} class implements the non-standard
  Bernardi-Raugel (BR) element that can be used to construct a stable
  velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
  element is an enriched version of the $Q_1^d$ element with added bubble
  functions on each edge (in 2d) or face (in 3d). It addresses the fact
  that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
  larger velocity space), and that the $Q_2^d\times Q_0$ combination is
  stable but converges with only first-order at the cost of the large
  number of velocity unknowns. The BR space is thus intermediate to the
  $Q_1^d$ and $Q_2^d$ spaces.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dedicated support for symbolic algebra, including algebra differentiation}
\label{subsec:sd}
To complement the automatic differentiation features in \dealii{}, this release sees
the first step towards integrating and supporting a highly performant computer algebra
system (CAS) via the SymEngine library \cite{symengine-web-page}.
The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces
to SymEngine and forms the basis of symbolic computations, offering a full set of
overloaded operators and a C++ style interface. This class offers the following
basic functionality:
\begin{itemize}
\item symbolic variable definition,
\item symbolic function definition,
\item expression creation using standard C++ syntax,
\item expression parsing,
\item comparison operations,
\item logical operations,
\item conditional expression construction,
\item differentiation,
\item substitution (partial and complete), and
\item serialization.
\end{itemize}
\dealii{} now also provides an extensive set of math operations, with a syntax mimicking that used
in the C++ standard library. Using the \texttt{Expression} class as a basis, we
have developed a set of functions that can be used to create \dealii{} \texttt{Tensor}s
and \texttt{SymmetricTensor}s of symbolic variables and symbolic functions. This
gives full symbolic tensor algebra support using the pre-existing \texttt{Tensor}
and \texttt{SymmetricTensor} classes and associated functions. We have
also implemented a set of utility functions with the following features:
\begin{itemize}
\item \texttt{differentiate} scalar expressions, tensor and symmetric tensors of expressions
with respect to other scalar expressions, tensor and symmetric tensors of expressions;
\item create symbolic substitution maps;
\item resolve explicit dependencies between expressions; and
\item perform scalar and tensor valued substitution (including conversion from symbolic to
real-valued scalars and tensors).
\end{itemize}

@@

GPU support was significantly extended for the current release:
\begin{itemize}
  \item The \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
    classes can now be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
  \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
    class has gained a second template argument \texttt{MemorySpace} which can
    either be \texttt{Host} or \texttt{CUDA}. In the latter case, the data
    resides in GPU memory. By default, the template parameter is
    \texttt{Host} and the behavior is unchanged compared to previous versions.
    When using CUDA, the ghost exchange can be performed either by first copying
    the relevant data to the host, performing MPI communication, and finally
    move the data to the device or, if CUDA-aware MPI is available, by
    performing MPI communication directly between GPUs.
  \item Constrained degrees of freedom: the matrix-free framework now
    supports constrained degrees of freedom. The implementation is based on
    the algorithms described in \cite{ljungkvist2017}. With this addition,
    both Dirichlet boundary conditions and the constraints arising from
    adaptively refined meshes can be imposed within the matrix-free
    framework.
    \textcolor{red}{%
    \sout{The only restriction is that for two-dimensional meshes the
    finite element degree must be odd. There is no such restriction in
    three dimensions.}
    \emph{%
    WB: Can we explain this restriction? It seems odd (pun intended)...
    BT: No, we cannot. I have spent at least two full weeks trying to
    understand why it doesn't work, i.e., gives the wrong result, but I
    don't know. We decided at the time that even with the current
    limitation this was a useful addition.
    MM: I suggest that we then simply not mention the limitation and treat
    it as a known bug. As it is currently worded it can be easily
    interpreted as a fundamental limitation, i.\,e. ``FE with even degree
    are not admissible.''}}

  \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
    matrix-free framework can scale to multiple GPUs by taking
    advantage of MPI. Each MPI process can only use one GPU and therefore, if
    multiple GPUs are available in one node, it is necessary to have as many
    ranks as there are GPUs. Using Nvidia Multi-Process Service (MPS), it is also possible
    for multiple processes to use the same GPU. This can be advantageous if the
    amount of work on one rank is not sufficient to fully utilize a GPU.
\end{itemize}

@@

\textcolor{red}{Missing: performance optimizations, new features, largest
runs now at XYZ (Kronbichler?)}

@@

\dealii can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{Griewank1996a,adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK \cite{lapack}
\item cuSOLVER \cite{cusolver}
\item cuSPARSE \cite{cusparse}
\item Gmsh \cite{geuzaine2009gmsh}
\item GSL \cite{gsl2016}
\item Ginkgo \cite{ginkgo-web-page}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{Burstedde2018,p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK \cite{slug}
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item SymEngine \cite{symengine-web-page}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

@@

\dealii is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
% updated 6/05/2019
  Giovanni Alzetta,
  Mathias Anselmann,
  Daniel Appel,
  Alexander Blank,
  Vishal Boddu,
  Benjamin Brands,
  Pi-Yueh Chuang,
  T. Conrad Clevenger,
  Sambit Das,
  Stefano Dominici,
  Nivesh Dommaraju,
  Niklas Fehn,
  Isuru Fernando,
  Andreas Fink,
  Rene Gassm{\"o}ller,
  Alexander Grayver,
  Joshua Hanophy,
  Logan Harbour,
  Daniel Jodlbauer,
  Stefan Kaessmair,
  Eldar Khattatov,
  Alexander Knieps,
  Uwe K{\"o}cher,
  Kurt Kremitzki,
  Dustin Kumor,
  Damien Lebrun-Grandie,
  Jonathan Matthews,
  Stefan Meggendorfer,
  Pratik Nayak,
  Lei Qiao,
  Ce Qin,
  Reza Rastak,
  Roland Richter,
  Alberto Sartori,
  Svenja Schoeder,
  Sebastian Stark,
  Antoni Vidal,
  Jiaxin Wang,
  Yuxiang Wang,
  Zhuoran Wang.
